A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-22 tex/codatadp.tex
more

Given the mapping $\Val$, the \rn{Acyclic} and \rn{Unique} rules can be stated as follows.
For acyclicity, if there exists an equivalence class $\ec{t}$
where $t$'s type is a datatype and
whose %class of  -- "whose" already refers to "equivalence class"
values $\Val \ec{t} = \MU x.\; u$ are cyclic,
as indicated by the assumption $x \in \FV( u )$,
then $\Ec$ is unsatisfiable.
For uniqueness, if there exist two equivalence classes $\ec{t}$ and $\ec{u}$
where $t$ and $u$'s type is a codatatype and
whose %classes of  -- "whose" already refers to "equivalence classes"
values $\Val \ec{t}$ and $\Val \ec{u}$ are $\alpha$-equivalent,
%for a capture-avoiding renaming of $\mu$-bound variables,
then $t \teq u$. This rule may seem too restrictive, since
$\MU x.\; \const{S}(x)$ and $\MU y.\; \const{S}(\const{S}(y))$ denote the same
value despite being $\alpha$-disequivalent, but the construction also works in
such cases, as illustrated by the following example.

@@

%We first state the following properties of $\Val^\AST$,
In the following lemma about $\Val^\AST$,
$\muvar( t )$ denotes $t$ if $t$ is a variable and ${\Varec{x}}$ if $t$ is of the form $\MU {\Varec{x}}.\; u$.

>>>>

Given the mapping $\Val$, the \rn{Acyclic} and \rn{Unique} rules can be stated as follows.
For acyclicity, if $\ec{t : \delta}$ is an equivalence class such that $\delta$ is a datatype and
whose %class of  -- "whose" already refers to "equivalence class"
values $\Val \ec{t} = \MU x.\; u$ are cyclic,
as indicated by the assumption $x \in \FV( u )$,
then $\Ec$ is unsatisfiable.
For uniqueness, if $\ec{t : \delta}$ and $\ec{u : \delta}$ are two equivalence classes such
that $\delta$ is a codatatype and
whose %classes of  -- "whose" already refers to "equivalence classes"
values $\Val \ec{t}$ and $\Val \ec{u}$ are $\alpha$-equivalent,
%for a capture-avoiding renaming of $\mu$-bound variables,
then $t \teq u$. This rule may seem too restrictive, since
$\MU x.\; \const{S}(x)$ and $\MU y.\; \const{S}(\const{S}(y))$ denote the same
value despite being $\alpha$-disequivalent, but the construction also works in
such cases, as illustrated by the following example.

@@

%We first state the following properties of $\Val^\AST$,
In the following lemma about $\Val^\AST$,
$\muvar( t ) = \begin{cases}
  x & \text{if $t = \MU {x}.\; u$} \\[-\jot]
  t & \text{otherwise ($t$ is a variable)}
\end{cases}$
